---
title: "Exploring U.S. Federal Holidays with assign4"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring U.S. Federal Holidays Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Load necessary packages and data
library(assign4)   # Load the package
library(dplyr)     # For data manipulation
library(janitor)   # For cleaning column names
library(lubridate) # For working with dates
library(ggplot2)   # For plotting
library(stringr)   # For string manipulation (str_extract, etc.)
```

## Introduction

The **assign4** package provides tools for analyzing and exploring U.S. Federal Holidays, using both established and proposed holiday datasets. This package combines static analysis (from Assignment 2) with an interactive Shiny app that allows users to explore holiday data dynamically.

The package offers functions for data cleaning, visualization, and interactive exploration of federal holidays, with a particular focus on:
- Historical trends in holiday creation.
- Thematic categorization of proposed holidays.
- Interactive exploration of holiday data.

This vignette will guide you through the core features of the package, including the data processing used to clean the datasets and examples of how to use the Shiny app.

## Data Processing and Cleaning

The data used in this package is derived from the **TidyTuesday** project (June 18, 2024). The dataset includes both **federal holidays** and **proposed federal holidays**. Below, we show the full code to clean and process these datasets, which are included in the package as `federal_holidays` and `proposed_federal_holidays`.

### Federal Holidays Dataset Processing

The **federal_holidays** dataset contains information about established U.S. federal holidays, including their names, establishment dates, and other details. Here's how the data was cleaned:

```{r federal-holidays-cleaning, eval=TRUE}
# Load the raw dataset (TidyTuesday data or local file in actual use)
# tuesdata <- tidytuesdayR::tt_load('2024-06-18')
# federal_holidays <- tuesdata$federal_holidays

# For the vignette, we'll assume the dataset is already included in the package
data(federal_holidays)

# Clean the dataset
federal_holidays_clean <- federal_holidays %>%
  janitor::clean_names() %>%
  mutate(
    year_established = as.integer(year_established),
    date_established = lubridate::mdy(date_established),
    details = stringr::str_remove_all(details, "\\[\\d+\\]")
  )

# View the first few rows of the cleaned dataset
head(federal_holidays_clean)
```

### Proposed Federal Holidays Dataset Processing

The **proposed_federal_holidays** dataset includes proposals for new U.S. federal holidays, along with details about the proposed dates and the reasons for the proposals.

```{r proposed-holidays-cleaning, eval=TRUE}
# Load the proposed federal holidays dataset
# proposed_federal_holidays <- tuesdata$proposed_federal_holidays

# For the vignette, we'll assume the dataset is already included in the package
data(proposed_federal_holidays)

# Clean the dataset
proposed_federal_holidays_clean <- proposed_federal_holidays %>%
  janitor::clean_names() %>%
  mutate(
    start_date = str_extract(date, "^[A-Za-z]+ \\d+"),
    end_date = str_extract(date, "\\d+$"),
    end_date = ifelse(is.na(end_date), start_date,
                      paste(str_extract(start_date, "^[A-Za-z]+"), end_date)),
    date_definition = stringr::str_remove_all(date_definition, "\\[\\d+\\]"),
    details = stringr::str_remove_all(details, "\\[\\d+\\]")
  ) %>%
  mutate(
    month = str_extract(start_date, "^[A-Za-z]+")  # Extract month from start date
  )

# View the first few rows of the cleaned dataset
head(proposed_federal_holidays_clean)
```

### Categorizing Proposed Holidays by Theme

To better understand the proposed holidays, we can categorize them by theme. For example, we can group holidays into categories such as **Cultural Recognition**, **Environmental**, **Political**, and **Patriotic**.

```{r proposed-holidays-categorization, eval=TRUE}
# Categorize proposed holidays by theme
proposed_category <- proposed_federal_holidays_clean %>%
  mutate(category = case_when(
    str_detect(details, "Indigenous|Native|Cesar Chavez|Rosa Parks|Harriet Tubman|Malcolm X|Susan B. Anthony") ~ "Cultural Recognition",
    str_detect(details, "Environment|Earth|Climate") ~ "Environmental",
    str_detect(details, "Election|Voting|Democracy|Voter|President|Congress|Act") ~ "Political",
    str_detect(details, "Flag|September 11|Remembrance|Veterans|Memorial|Patriotic") ~ "Patriotic",
    TRUE ~ "Other"
  ))

# Count holidays by category
holiday_categories <- proposed_category %>%
  count(category)

# Display the categories
holiday_categories %>%
  knitr::kable(col.names = c("Holiday Category", "Count"))
```

## Key Results from Data Analysis

The timeline of U.S. federal holidays reveals that many holidays were introduced during times of significant social or political change. For example, holidays like **Labor Day** (1894) arose during the industrial labor movement, while **Martin Luther King Jr. Day** (1983) highlights the nation's growing recognition of civil rights.

### Table: Recent Federal Holidays

This table shows the most recent U.S. federal holidays:

```{r table-recent-holidays, eval=TRUE}
federal_holidays_clean %>%
  arrange(desc(year_established)) %>%
  select(official_name, year_established) %>%
  head(5) %>%
  knitr::kable(col.names = c("Holiday Name", "Year Established"))
```

The most recent addition to the federal holiday calendar is **Juneteenth National Independence Day**, established in 2021.

## Using the Shiny App

One of the key features of **assign4** is the Shiny app, which allows users to explore U.S. federal holidays interactively.

To launch the Shiny app, use the following command:

```{r, eval = FALSE}
# Launch the Shiny app
launch_shiny_app()
```

### Features of the Shiny App:

1. **2024 Federal Holidays**: View a table of the official U.S. federal holidays for 2024, with the ability to filter by month.
2. **Proposed Federal Holidays**: Explore proposed holidays, categorized by theme (e.g., cultural recognition, environmental, political).
3. **Analytics**: Visualize the timeline of holiday establishment and analyze the distribution of proposed holidays by month.

## Example: Holiday Trends

The following plot shows the historical timeline of U.S. federal holidays, revealing how holidays cluster around key social and political movements.

```{r fig-timeline-established-holidays, eval=TRUE, fig.cap="Establishment of U.S. Federal Holidays", fig.align='center'}
ggplot(federal_holidays_clean, aes(x = year_established, y = reorder(official_name, year_established))) +
  geom_point(color = "blue", size = 3) +
  labs(
    title = "Establishment of U.S. Federal Holidays",
    x = "Year Established",
    y = "Holiday",
    caption = "Source: U.S. Federal Holidays Dataset"
  ) +
  theme_minimal()
```

This plot shows that significant moments in U.S. history, such as the labor movement and civil rights movement, are often marked by the establishment of federal holidays.

## Conclusion

The **assign4** package offers a comprehensive set of tools for exploring U.S. federal holidays. By combining static analysis with interactive features in the Shiny app, this package provides users with a deeper understanding of how holidays reflect broader social and political movements in the U.S.

For more information, refer to the [package documentation](https://your-pkgdown-site-link) and the interactive [Shiny app](https://your-shiny-app-link).
