---
title: "Exploring U.S. Federal Holidays Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring U.S. Federal Holidays Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### Introduction

This vignette provides an overview of the U.S. Federal Holidays package, exploring its features and how to use it for analyzing trends in federal holidays and proposals for new ones.

### Loading Data

We first load the data using the tidytuesdayR package:

```{r}
library(tidytuesdayR)
tuesdata <- tidytuesdayR::tt_load('2024-06-18')
federal_holidays <- tuesdata$federal_holidays
proposed_federal_holidays <- tuesdata$proposed_federal_holidays
```

### Data Cleaning and Preparation

We clean and prepare the data using the tidyverse and janitor packages:

```{r}
library(tidyverse)
library(janitor)

federal_holidays <- federal_holidays %>%
  janitor::clean_names() %>%
  mutate(year_established = as.integer(year_established))
```

### Analyzing Established Holidays

Visualize the timeline of U.S. federal holidays:

```{r}
library(ggplot2)

ggplot(federal_holidays, aes(x = year_established, y = reorder(official_name, year_established))) +
  geom_point(color = "blue", size = 3) +
  labs(
    title = "Establishment of U.S. Federal Holidays",
    x = "Year Established",
    y = "Holiday"
  ) +
  theme_minimal()
```

### Analyzing Proposed Holidays

Explore the distribution of proposed holidays:

```{r}
proposed_holidays <- proposed_federal_holidays %>%
  mutate(month = lubridate::month(date, label = TRUE)) %>%
  count(month)

ggplot(proposed_holidays, aes(x = month, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Distribution of Proposed Federal Holidays by Month",
       x = "Month", y = "Count") +
  theme_minimal()
```

### Conclusion

The vignette demonstrates how to use the package to explore the evolution and proposals of U.S. federal holidays, reflecting cultural and societal trends.

#### Steps to Follow:

1. **Create and Edit the Vignette**: Use `usethis::use_vignette("explore-your-package")` to create the vignette file. Then, edit it with the content above.

2. **Build Vignettes**: Run `devtools::build_vignettes()` to build the vignette.

3. **Check Vignettes**: Use `browseVignettes(package = "yourpackage")` to view the vignette and ensure it displays correctly.

This structure provides a comprehensive guide for users to explore your package's functionalities.






